---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Context

Buccaneer is an R package for estimating competition and coexistence metrics 
    through deep time using fossil data. It lets you explore how competition 
    changes across different biological levels (clades, species, and assemblages) 
    and spatial scales (local and regional).

To get started, you only need a few core inputs: species longevities, fossil 
    occurrence records with spatial information, and species traits. Buccaneer’s
    main functions then generate flexible, customizable time series of competition
    and co-occurrence metrics that can be tailored to a wide range of 
    evolutionary and paleoecological questions.
    
# Download 

To install the development version of this package the user must type:

```{.r}
# install.packages("devtools")
devtools::install_github("GabrielNakamura/Buccaneer", ref = "main")
```

# Help page

```{r,eval=FALSE}
help(package = "Buccaneer")
```


