---
title: "Untitled"
author: "Gabriel Nakamura"
date: "2025-07-15"
output: html_document
---


# Canidae data 

```{r}
devtools::load_all()

load(here::here("inst", "extdata", "df_can.Rdata"))
load(here::here("inst", "extdata", "longevities.Rdata"))

library(tidyr)
library(dplyr)
library(ggplot2)
```

Computing mean longevities

```{r}
list_df_TS_TE <- 
  lapply(longs, function(x){
  data.frame(species = rownames(x), TS = x$TS, TE = x$TE)
})

df_longevities <- do.call(rbind, list_df_TS_TE)

# median values for TS and TE
df_longevities2 <- 
  df_longevities |> 
  group_by(species) |> 
  mutate(TS = median(TS), 
         TE = median(TE)) |> 
  distinct(species, .keep_all = TRUE) 

```

Getting occurrences

```{r}

df_occ <- df_can

```

# trait data

```{r}
df_traits <- read.csv(here::here("inst", "extdata", "final_data.csv"), sep = ",")
df_traits2 <- 
  df_traits |> 
  select(LD1, log_mass)

dist_traits <- dist(df_traits2)
```


# computing coexisting metrics


```{r}
res_clade_distance_site <- 
  clade_site_distance(df.TS.TE = df_longevities2, 
                      df.occ = df_occ, 
                      time.slice = 0.1, 
                      dist.trait = dist_traits, 
                      nearest.taxon = "all",
                      round.digits = 1,
                      species = "species",
                      TS = "TS",
                      TE = "TE", 
                      Max.age = "max", 
                      Min.age = "min",
                      site = "site")
```

# computing null model

Computing null model for site scale considering mean distance for all coexisting
    species in a site 

First computing site occurrences


```{r}

# computing species in each timeslice
time_slice <- 0.1
seq_interval <- seq(from = max(df_longevities2[, "TS"]), to = min(df_longevities2[, "TE"]), by = -time_slice)
seq_interval <- c(round(seq_interval, digits = 1))

spp_slice <- 
  calc_spp_slice(df.TS.TE = df_longevities2,
                 time.slice = 0.1,
                 round.digits = 1, 
                 seq.interval = seq_interval,
                 species = "species",
                 TS = "TS",
                 TE = "TE")

list_composition <- 
  comp_site_occurrence (spp_slice = spp_slice,
                        df.occ = df_occ,
                        species = "species",
                        Max.age = "max",
                        Min.age = "min",
                        site = "site")

```

Now computing null models for sites

```{r}

# first get the complete distance matrix

matrix_dist_complete <- as.matrix(dist_traits)
rownames(matrix_dist_complete) <- df_longevities2$species
colnames(matrix_dist_complete) <- df_longevities2$species

df_null_site_all <- 
  calc_null_model(list_occurrence = list_composition, 
                  dist_matrix_trait = matrix_dist_complete, 
                  nearest.taxon = "all",
                  nperm = 1000)
```

calculating z scores

```{r}
vec_mean_null <- apply(df_null_site_all, 2, function(x) mean(x))
sd_mean_null <- apply(df_null_site_all, 2, function(x) sd(x))

# p values

df_mean_obs_null <- data.frame(mean.null = vec_mean_null, sd.mean.null = sd_mean_null, obs.dist = res_clade_distance_site$mean.distance)

df_mean_obs_null2 <- 
  df_mean_obs_null |> 
  mutate(z.values = ((obs.dist - mean.null) / (sd.mean.null))) |> 
  mutate(dist.values = abs(obs.dist - mean.null)) 

p_values <- 
  unlist(lapply(1:nrow(df_mean_obs_null2), function(x){
    mean(abs(df_null_site_all[, x] - df_mean_obs_null2$mean.null[x]) >= df_mean_obs_null2$dist.values[x])
  }))

# complete data frame for plotting results
df_mean_obs_null3 <- data.frame(df_mean_obs_null2, p.values = p_values, time.slice = seq_interval)





```


# Plotting site results

```{r}
# Example data setup
# df <- read.csv("your_data.csv")  # if loading from a file

# Ensure variables are numeric
df_mean_obs_null3$time.slice <- as.numeric(df_mean_obs_null3$time.slice)
df_mean_obs_null3$z.values <- as.numeric(df_mean_obs_null3$z.values)
df_mean_obs_null3$p.values <- as.numeric(df_mean_obs_null3$p.values)


df_mean_obs_null3$significant <- df_mean_obs_null3$p.values <= 0.05

library(ggplot2)

plot_site_mpd <- 
  ggplot(df_mean_obs_null3, aes(x = time.slice, y = z.values)) +
  geom_line(color = "grey50", alpha = 0.4) +  # Transparent line
  geom_smooth(method = "loess", se = TRUE, color = "#61c2e2", size = 3) +
  geom_point(aes(color = significant), size = 4, alpha = 0.4) +  # Transparent points
  scale_color_manual(values = c("TRUE" = "#e12a44", "FALSE" = "#9fb493")) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  scale_x_reverse() +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    legend.position = "none",
    axis.line = element_line(color = "black"),
    axis.ticks = element_line(color = "black"), 
    axis.text.x = element_text(size = 26), 
    axis.text.y = element_text(size = 26), 
    axis.title.x = element_text(size = 26),
    axis.title.y = element_text(size = 26)
  ) +
  labs(x = "Time", y = "SES - mean pairwise distance")

ggsave(here::here("inst", "extdata", "figures_canidae", "plot_site_sesMPD.png"),
       plot_site_mpd, 
       width = 10, 
       height = 10, 
       dpi = 800,
       bg = "white")

```

# For nearest neighbour

Computing mean longevities

```{r}
list_df_TS_TE <- 
  lapply(longs, function(x){
  data.frame(species = rownames(x), TS = x$TS, TE = x$TE)
})

df_longevities <- do.call(rbind, list_df_TS_TE)

# median values for TS and TE
df_longevities2 <- 
  df_longevities |> 
  group_by(species) |> 
  mutate(TS = median(TS), 
         TE = median(TE)) |> 
  distinct(species, .keep_all = TRUE) 

```

Getting occurrences

```{r}

df_occ <- df_can

```

# trait data

```{r}
df_traits <- read.csv(here::here("inst", "extdata", "final_data.csv"), sep = ",")
df_traits2 <- 
  df_traits |> 
  select(LD1, log_mass)

dist_traits <- dist(df_traits2)
```


# computing coexisting metrics


```{r}
res_clade_distance_site_nearest <- 
  clade_site_distance(df.TS.TE = df_longevities2, 
                      df.occ = df_occ, 
                      time.slice = 0.1, 
                      dist.trait = dist_traits, 
                      nearest.taxon = 1,
                      round.digits = 1,
                      species = "species",
                      TS = "TS",
                      TE = "TE", 
                      Max.age = "max", 
                      Min.age = "min",
                      site = "site")
```

# computing null model

Computing null model for site scale considering mean distance for all coexisting
    species in a site 

First computing site occurrences


```{r}

# computing species in each timeslice
time_slice <- 0.1
seq_interval <- seq(from = max(df_longevities2[, "TS"]), to = min(df_longevities2[, "TE"]), by = -time_slice)
seq_interval <- c(round(seq_interval, digits = 1))

spp_slice <- 
  calc_spp_slice(df.TS.TE = df_longevities2,
                 time.slice = 0.1,
                 round.digits = 1, 
                 seq.interval = seq_interval,
                 species = "species",
                 TS = "TS",
                 TE = "TE")

list_composition <- 
  comp_site_occurrence (spp_slice = spp_slice,
                        df.occ = df_occ,
                        species = "species",
                        Max.age = "max",
                        Min.age = "min",
                        site = "site")

```

Now computing null models for sites

```{r}

# first get the complete distance matrix

matrix_dist_complete <- as.matrix(dist_traits)
rownames(matrix_dist_complete) <- df_longevities2$species
colnames(matrix_dist_complete) <- df_longevities2$species

df_null_site_all_nearest <- 
  calc_null_model(list_occurrence = list_composition, 
                  dist_matrix_trait = matrix_dist_complete, 
                  nearest.taxon = 1,
                  nperm = 1000)
```

calculating z scores

```{r}
vec_mean_null_nearest <- apply(df_null_site_all_nearest, 2, function(x) mean(x))
sd_mean_null_nearest <- apply(df_null_site_all_nearest, 2, function(x) sd(x))

# p values

df_mean_obs_null_nearest <- data.frame(mean.null = vec_mean_null_nearest, sd.mean.null = sd_mean_null_nearest, obs.dist = res_clade_distance_site_nearest$mean.distance)

df_mean_obs_null2_nearest <- 
  df_mean_obs_null_nearest |> 
  mutate(z.values = ((obs.dist - mean.null) / (sd.mean.null))) |> 
  mutate(dist.values = abs(obs.dist - mean.null)) 

p_values_nearest <- 
  unlist(lapply(1:nrow(df_mean_obs_null2_nearest), function(x){
    mean(abs(df_null_site_all_nearest[, x] - df_mean_obs_null2_nearest$mean.null[x]) >= df_mean_obs_null2_nearest$dist.values[x])
  }))

# complete data frame for plotting results
df_mean_obs_null3_nearest <- data.frame(df_mean_obs_null2_nearest, p.values = p_values_nearest, time.slice = seq_interval)





```


# Plotting site results

```{r}
# Example data setup
# df <- read.csv("your_data.csv")  # if loading from a file

# Ensure variables are numeric
df_mean_obs_null3_nearest$time.slice <- as.numeric(df_mean_obs_null3_nearest$time.slice)
df_mean_obs_null3_nearest$z.values <- as.numeric(df_mean_obs_null3_nearest$z.values)
df_mean_obs_null3_nearest$p.values <- as.numeric(df_mean_obs_null3_nearest$p.values)


df_mean_obs_null3_nearest$significant <- df_mean_obs_null3_nearest$p.values <= 0.05

library(ggplot2)

plot_site_mpd_nearest <- 
  ggplot(df_mean_obs_null3_nearest, aes(x = time.slice, y = z.values)) +
  geom_line(color = "grey50", alpha = 0.4) +  # Transparent line
  geom_smooth(method = "loess", se = TRUE, color = "#61c2e2", size = 3) +
  geom_point(aes(color = significant), size = 4, alpha = 0.4) +  # Transparent points
  scale_color_manual(values = c("TRUE" = "#e12a44", "FALSE" = "#9fb493")) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  scale_x_reverse() +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    legend.position = "none",
    axis.line = element_line(color = "black"),
    axis.ticks = element_line(color = "black"), 
    axis.text.x = element_text(size = 26), 
    axis.text.y = element_text(size = 26), 
    axis.title.x = element_text(size = 26),
    axis.title.y = element_text(size = 26)
  ) +
  labs(x = "Time", y = "SES - mean nearest distance")

ggsave(here::here("inst", "extdata", "figures_canidae", "plot_site_sesMPD_nearest.png"),
       plot_site_mpd_nearest, 
       width = 10, 
       height = 10, 
       dpi = 800,
       bg = "white")

```


# plotting raw data for mean distance 

```{r}
res_clade_distance_site

plot_site_mpd_raw <- 
  ggplot(res_clade_distance_site, aes(x = time.slice, y = mean.distance)) +
  geom_line(color = "grey50", alpha = 0.4) +  # Transparent line
  geom_smooth(method = "loess", se = TRUE, color = "#61c2e2", size = 3) +
  geom_point(aes(color = "#9fb493"), size = 3, alpha = 0.4) +  # Transparent points
  scale_x_reverse() +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    legend.position = "none",
    axis.line = element_line(color = "black"),
    axis.ticks = element_line(color = "black"), 
    axis.text.x = element_text(size = 26), 
    axis.text.y = element_text(size = 26), 
    axis.title.x = element_text(size = 26),
    axis.title.y = element_text(size = 26)
  ) +
  labs(x = "Time", y = "Mean Nearest Distance")


ggsave(here::here("inst", "extdata", "figures_canidae", "plot_site_MPD_raw.png"),
       plot_site_mpd_raw, 
       width = 10, 
       height = 10, 
       dpi = 800,
       bg = "white")

```


# Additional graphs for raw data

```{r}

plot_site_mnnd_raw <- 
  ggplot(res_clade_distance_site_nearest, aes(x = time.slice, y = mean.distance)) +
  geom_line(color = "grey50", alpha = 0.4) +  # Transparent line
  geom_smooth(method = "loess", se = TRUE, color = "#9fb493", size = 3) +
  scale_x_reverse() +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    legend.position = "none",
    axis.line = element_line(color = "black"),
    axis.ticks = element_line(color = "black"), 
    axis.text.x = element_text(size = 26), 
    axis.text.y = element_text(size = 26), 
    axis.title.x = element_text(size = 26),
    axis.title.y = element_text(size = 26)
  ) +
  labs(x = "Time", y = "Mean Nearest Distance")


ggsave(here::here("inst", "extdata", "figures_canidae", "plot_site_mnnd_raw.png"),
       plot_site_mnnd_raw, 
       width = 10, 
       height = 10, 
       dpi = 800,
       bg = "white")
```

