---
title: "Time series metrics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{time-series-metrics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ggplot2)
library(dplyr)
```

# Background

Here we will use the same data processed in the article xXXXX.


# Loading data and packages

```{r}
data("df_TS_TE_faurby")
data("df_TS_TE_mass")
data("df_occ_faurby")

```

## Clade level analysis 

Calculating time series at clade level 

```{r}
regional_richness <- 
  regional_clade_richness(df.TS.TE = df_TS_TE_faurby,
                          time.slice = 0.1,
                          round.digits = 1,
                          species = "species",
                          TS = "TS",
                          TE = "TE")
```

Plotting the results

```{r, out.width="80%", out.height="80%", dpi=300, fig.retina=2}
regional_richness |> 
  mutate(time.slice = as.numeric(time.slice)) |> 
  ggplot(aes(x = time.slice, y = richness)) +
  geom_smooth(se = TRUE, method = "loess", size = 1) +  # Line thickness for better visibility
  geom_point(aes(x = time.slice, y = richness)) +
  labs(title = "",
       x = "Tempo",
       y = "Riqueza") +
  scale_x_continuous(breaks = seq(max(as.numeric(regional_richness$time.slice)), 0, by = -10)) +
  scale_x_reverse() +
  theme_minimal() +
  theme(legend.position = "none")# Use a clean theme
```

### regional clade mpd

Calculating regional clade level mpd with body size. First with all species

```{r}

regional_mpd <- 
  clade_regional_mpd(df.TS.TE = df_TS_TE_mass, 
                     time.slice = 0.1,
                     trait = "mean.size",
                     round.digits = 1,
                     species = "species",
                     TS = "TS",
                     TE = "TE",
                     compute.ses.mpd = TRUE)
```

Calculating comparing mpd between and within groups

```{r}
regional_mpd_between <- 
  clade_regional_mpd(df.TS.TE = df_TS_TE_mass, 
                     time.slice = 0.1,
                     trait = "mean.size",
                     round.digits = 1,
                     species = "species",
                     TS = "TS",
                     TE = "TE",
                     compute.ses.mpd = TRUE,
                     null.model = "taxa.labels",
                     runs = 999, 
                     group = "group", 
                     group.focal.compare = c("Caniformia", "Feliformia"),
                     type.comparison = "between")
```
Calculating within species considering only Caniformia as the focal group

```{r}
regional_mpd_within <- 
  clade_regional_mpd(df.TS.TE = df_TS_TE_mass, 
                     time.slice = 0.1,
                     trait = "mean.size",
                     round.digits = 1,
                     species = "species",
                     TS = "TS",
                     TE = "TE",
                     compute.ses.mpd = TRUE,
                     null.model = "taxa.labels",
                     runs = 999, 
                     group = "group", 
                     group.focal.compare = c("Caniformia", "Feliformia"),
                     type.comparison = "within")
```

Joining all results to plot in one graphic

```{r}
all_mpd_regional <- rbind(regional_mpd, regional_mpd_between, regional_mpd_within)
all_mpd_regional2 <- data.frame(all_mpd_regional, group_res = rep(c("All", "Between", "Within"), each = nrow(regional_mpd_within)))

```



Plotting mpd results

```{r, out.width="80%", out.height="80%", dpi=300, fig.retina=2}
all_mpd_regional2 |> 
  ggplot(aes(x = time.slice, y = ses.mpd, color = group_res)) +
  geom_point(aes(x = time.slice, y = ses.mpd)) +
  geom_smooth(se = TRUE, method = "loess", size = 1) +  # Line thickness for better visibility
  labs(title = "",
       x = "Time",
       y = "Mean Pairwise Distance (ses)") +
  scale_x_continuous(breaks = seq(40, 0, by = -10)) +
  xlim(40, 0) +
  theme_minimal() +
  theme(legend.position = "bottom")# Use a clean theme
```

### Clade site richness

```{r}
site_richness <- 
  clade_site_richness(df.TS.TE = df_TS_TE_faurby, 
                      df.occ = df_occ_faurby, 
                      time.slice = 0.1, 
                      round.digits = 1,
                      species = "species",
                      TS = "TS",
                      TE = "TE")
```
Plotting the results

```{r}
site_richness |> 
  ggplot(aes(x = as.numeric(time.slice), y = mean.richness.site)) +
  geom_point(aes(x = as.numeric(time.slice), y = mean.richness.site)) +
  geom_smooth(se = TRUE, method = "loess", size = 1) +  # Line thickness for better visibility
  labs(title = "",
       x = "Time",
       y = "Mean site richness") +
  scale_x_continuous(breaks = seq(max(as.numeric(site_richness$time.slice)), 0, by = -10)) +
  scale_x_reverse() +
  theme_minimal() +
  theme(legend.position = "none")# Use a clean theme
```


## Individual species analysis

### Individual species regional coexistence

```{r}
indiv_species_coex <- 
  IndivSpecies_regional_richness(df.TS.TE = df_TS_TE_faurby, 
                                 time.slice = 0.1,
                                 round.digits = 1,
                                 species = "species",
                                 TS = "TS",
                                 TE = "TE")

```
This output contain two data frames, one with all individual coexistences per species in all time slices, and another containing the mean species coexistence considering all time slices

```{r}
df_mean_individual_species <- indiv_species_coex$mean_species_coexistence
df_indiv_species_long <- indiv_species_coex$df_IndivSpp_coexist
```


We can plot the results of all time mean species coexistence for three species 

```{r, out.width="80%", out.height="80%", dpi=300, fig.retina=2}
df_indiv_species_long |> 
  filter(species == "Canis" | species == "Ursus" | species == "Ursavus") |> 
  ggplot(aes(x = species, y = n.coexistence, fill = species)) +
  geom_violin(trim = FALSE) +  # Line thickness for better visibility
  geom_point(aes(x = species, y = n.coexistence, fill = species), shape = 21) +
  labs(title = "",
       x = "Espécies",
       y = "Coexistência média") +
  theme_minimal() +
  theme(legend.position = "none", 
        panel.grid.major = element_line(linetype = "dashed"))# Use a clean theme
```

We can also plot the time series for each species

```{r, out.width="80%", out.height="80%", dpi=300, fig.retina=2}
df_indiv_species_long |> 
  filter(species == "Canis" | species == "Ursavus" | species == "Ursus" | species == "Panthera") |> 
  ggplot(aes(x = time.slice, y = n.coexistence, color = species, group = species)) +  # Line thickness for better visibility
  geom_smooth(se = TRUE, method = "loess", size = 1) +
  geom_point(aes(x = time.slice, y = n.coexistence, color = species), shape = 21) +
  facet_wrap(~species) +
  labs(title = "",
       x = "Time",
       y = "number of coexistences by species") +
  theme_minimal()  +
  scale_x_continuous(breaks = seq(max(df_indiv_species_long$time.slice), 0, by = -10)) +
  scale_x_reverse() +
  theme(legend.position = "none")# Use a clean theme
```


### Individual species regional mpd

We can also plot individual species mpd for all species coexisting in regional context.
    For this we will also use data on body mass for all species. There are different ways 
    in which mpd can be calculated. We can calculate it using a single trait or a combination
    of traits. The later require a distance trait matrix (passed to `dist` argument)
    containing the pairwise distances among species.
    
Here we show how to calculate mpd using a single trait, in this case body mass.


```{r}

df_TS_TE_mass2 <- 
  df_TS_TE_mass |> 
  filter(across(c(TS, TE), ~ !is.na(.)))

indiv_regional_species_mpd <-
  IndivSpecies_regional_mpd(df.TS.TE = df_TS_TE_mass, 
                            time.slice = 0.1, 
                            trait = "mean.size",
                            round.digits = 1, 
                            species = "species",
                            TS = "TS",
                            TE = "TE")

df_indiv_species_mpd_long <- indiv_regional_species_mpd$df_IndivSpp_mpd
```

Plotting the results for individual species mpd 

```{r, out.width="80%", out.height="80%", dpi=300, fig.retina=2}
df_indiv_species_mpd_long |> 
  filter(species == "Canis" | species == "Ursavus" | species == "Ursus" | species == "Panthera") |> 
  ggplot(aes(x = time.slice, y = mpd, color = species, group = species)) +  # Line thickness for better visibility
  geom_point(aes(x = time.slice, y = mpd, color = species), shape = 21) +
  geom_smooth(se = TRUE, method = "loess", size = 1) +
  facet_wrap(~species) +
  labs(title = "",
       x = "Tempo",
       y = "Distância média (mpd)") +
  theme_minimal()  +
  scale_x_continuous(breaks = seq(max(df_indiv_species_mpd_long$time.slice), 0, by = -10)) +
  scale_x_reverse() +
  theme(legend.position = "none")# Use a clean theme
```

#### Individual species regional mpd with varying distances and multiple traits

In order to calculate the mpd for multiple traits the user must inform an object
   of class dist containing all pairwise distances among species. The function can
   handle distances in different ways. It can use all pairwise distances to obtain
   the mpd, or the user can inform throught the argument `nearest.taxon` the 
   number of nearest species that will be used to calculate the mpd. The default is 
   `all` which is the same as calculating the mpd considering all species co-occurring 
   in a time slice.


### Individual species site coexistence

Here we will perform the same calculations but using site criteria for individual species coexistence

```{r}
indiv_species_site_coex <- 
  IndivSpec_site_richness(df.TS.TE = df_TS_TE_faurby,
                          df.occ = df_occ_faurby, 
                          time.slice = 0.1,
                          round.digits = 1,
                          species = "species",
                          TS = "TS",
                          TE = "TE",
                          Max.age = "Max.age",
                          Min.age = "Min.age",
                          site = "site")

indiv_species_site_coex2 <- 
  indiv_species_site_coex |> 
  mutate(time.slice = as.numeric(time.slice))


```

plotting species coexistence considering site coexistence

```{r, out.width="80%", out.height="80%", dpi=300, fig.retina=2}
indiv_species_site_coex2 |> 
  filter(species == "Alopecocyon" | species == "Simocyon" | species == "Prepoecilogale" | species == "Vormela" | species == "Hesperocyon") |> 
  ggplot(aes(x = time.slice, y = mean.per.spp, color = species, group = species)) +  # Line thickness for better visibility
  geom_smooth(se = TRUE, method = "loess", size = 1) +
  facet_wrap(~species) +
  labs(title = "",
       x = "Time",
       y = "mean coexistence") +
  theme_minimal()  +
  scale_x_continuous(breaks = seq(max(indiv_species_site_coex2$time.slice), 0, by = -10)) +
  scale_x_reverse() +
  theme(legend.position = "none")
```

### Individual species site mpd 

We can calculate mean pairwise distances of individual species coexistence metrics based on site co-occurrence

```{r}

df_TS_TE_mass3 <- 
  df_TS_TE_mass2 |> 
  rename(species = "Genus")


indiv_species_site_mpd <- 
  IndivSpec_site_mpd(df.TS.TE = df_TS_TE_mass3, 
                     df.occ = df_occ_faurby, 
                     time.slice = 0.1, 
                     trait = "mean.size",
                     round.digits = 1,
                     species = "species",
                     TS = "TS",
                     TE = "TE",
                     Max.age = "Max.age",
                     Min.age = "Min.age",
                     site = "site")
```


### Individual reach coexistence 

We will calculate individual mean coexistence using reach criteria

```{r}

indiv_species_reach_coex <- 
  IndivSpec_reach_richness(df.TS.TE = df_TS_TE_faurby,
                           df.occ = df_occ_faurby,
                           time.slice = 0.1, 
                           round.digits = 1,
                           species = "species",
                           TS = "TS",
                           TE = "TE",
                           Max.age = "Max.age",
                           Min.age = "Min.age",
                           lat = "lat",
                           lon = "lng")

```

plotting individual species coexistence with reach criteria

```{r, out.width="80%", out.height="80%", dpi=300, fig.retina=2}

indiv_species_reach_coex2 <- 
  indiv_species_reach_coex |> 
  mutate(time.slice = as.numeric(time.slice))

indiv_species_reach_coex2 |> 
  filter(species == "Ursus" | species == "Simocyon" | species == "Plionarctos" | species == "Vormela" | species == "Hesperocyon") |> 
  ggplot(aes(x = time.slice, y = n.coexistence, color = species, group = species)) +  # Line thickness for better visibility
  geom_smooth(se = TRUE, method = "loess", size = 1) +
  facet_wrap(~species) +
  labs(title = "",
       x = "Time",
       y = "mean coexistence") +
  theme_minimal()  +
  scale_x_continuous(breaks = seq(max(indiv_species_site_coex2$time.slice), 0, by = -10)) +
  scale_x_reverse() +
  theme(legend.position = "none")
```


### Individual reach coexistence mpd

```{r}

```

