---
title: "Time series metrics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{time-series-metrics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(buccaneer)
```

# Background

Here we will use the same data processed in the article xXXXX.


# Loading data and packages

```{r}
data("df_TS_TE_faurby")
data("df_TS_TE_mass")
data("df_occ_faurby")

```


## Individual species analysis

### Individual species regional coexistence

```{r}
indiv_species_coex <- 
  IndivSpecies_regional_richness(df.TS.TE = df_TS_TE_faurby, 
                                 time.slice = 0.1,
                                 round.digits = 1,
                                 species = "species",
                                 TS = "TS",
                                 TE = "TE")

```
This output contain two data frames, one with all individual coexistences per species in all time slices, and another containing the mean species coexistence considering all time slices

```{r}
df_mean_individual_species <- indiv_species_coex$mean_species_coexistence
df_indiv_species_long <- indiv_species_coex$df_IndivSpp_coexist
```


We can plot the results of all time mean species coexistence for three species 

```{r}
df_indiv_species_long |> 
  filter(species == "Alopecocyon" | species == "Protursus" | species == "Simocyon") |> 
  ggplot(aes(x = species, y = n.coexistence, fill = species)) +
  geom_violin(trim = FALSE) +  # Line thickness for better visibility
  labs(title = "",
       x = "Species",
       y = "mean coexistences by species") +
  theme_minimal() +
  theme(legend.position = "none")# Use a clean theme
```

We can also plot the time series for each species

```{r}
df_indiv_species_long |> 
  filter(species == "Alopecocyon" | species == "Protursus" | species == "Simocyon") |> 
  ggplot(aes(x = time.slice, y = n.coexistence, color = species, group = species)) +  # Line thickness for better visibility
  geom_smooth(se = TRUE, method = "loess", size = 1) +
  facet_wrap(~species) +
  labs(title = "",
       x = "Time",
       y = "number of coexistences by species") +
  theme_minimal()  +
  scale_x_continuous(breaks = seq(max(df_indiv_species_long$time.slice), 0, by = -10)) +
  scale_x_reverse() +
  theme(legend.position = "none")# Use a clean theme
```


### Individual species regional mpd

We can also plot individual species mpd for all species coexisting in regional context.
For this we will also use data on body mass for all species


```{r}

df_TS_TE_mass2 <- 
  df_TS_TE_mass |> 
  filter(across(c(TS, TE), ~ !is.na(.)))

indiv_regional_species_mpd <-
  IndivSpecies_regional_mpd(df.TS.TE = df_TS_TE_mass2, 
                            time.slice = 0.1, 
                            trait = "mean.size",
                            round.digits = 1, 
                            species = "Genus",
                            TS = "TS",
                            TE = "TE")

df_indiv_species_mpd_long<- indiv_regional_species_mpd$df_IndivSpp_mpd
```

Plotting the results for individual species mpd from the object

```{r}
df_indiv_species_mpd_long |> 
  filter(species == "Alopecocyon" | species == "Simocyon" | species == "Prepoecilogale" | species == "Vormela" | species == "Hesperocyon") |> 
  ggplot(aes(x = time.slice, y = mpd, color = species, group = species)) +  # Line thickness for better visibility
  geom_smooth(se = TRUE, method = "loess", size = 1) +
  facet_wrap(~species) +
  labs(title = "",
       x = "Time",
       y = "mean mpd") +
  theme_minimal()  +
  scale_x_continuous(breaks = seq(max(df_indiv_species_mpd_long$time.slice), 0, by = -10)) +
  scale_x_reverse() +
  theme(legend.position = "none")# Use a clean theme
```




### Individual species site coexistence

Here we will perform the same calculations but using site criteria for individual species coexistence

```{r}
indiv_species_site_coex <- 
  IndivSpec_site_richness(df.TS.TE = df_TS_TE_faurby,
                          df.occ = df_occ_faurby, 
                          time.slice = 0.1,
                          round.digits = 1,
                          species = "species",
                          TS = "TS",
                          TE = "TE",
                          Max.age = "Max.age",
                          Min.age = "Min.age",
                          site = "site")

indiv_species_site_coex2 <- 
  indiv_species_site_coex |> 
  mutate(time.slice = as.numeric(time.slice))


```

plotting species coexistence considering site coexistence

```{r}
indiv_species_site_coex2 |> 
  filter(species == "Alopecocyon" | species == "Simocyon" | species == "Prepoecilogale" | species == "Vormela" | species == "Hesperocyon") |> 
  ggplot(aes(x = time.slice, y = mean.per.spp, color = species, group = species)) +  # Line thickness for better visibility
  geom_smooth(se = TRUE, method = "loess", size = 1) +
  facet_wrap(~species) +
  labs(title = "",
       x = "Time",
       y = "mean coexistence") +
  theme_minimal()  +
  scale_x_continuous(breaks = seq(max(indiv_species_site_coex2$time.slice), 0, by = -10)) +
  scale_x_reverse() +
  theme(legend.position = "none")
```

### Individual species site mpd 

We can calculate mean pairwise distances of individual species coexistence metrics based on site co-occurrence

```{r}

df_TS_TE_mass3 <- 
  df_TS_TE_mass2 |> 
  rename(species = "Genus")


indiv_species_site_mpd <- 
  IndivSpec_site_mpd(df.TS.TE = df_TS_TE_mass3, 
                     df.occ = df_occ_faurby, 
                     time.slice = 0.1, 
                     trait = "mean.size",
                     round.digits = 1,
                     species = "species",
                     TS = "TS",
                     TE = "TE",
                     Max.age = "Max.age",
                     Min.age = "Min.age",
                     site = "site")
```


### Individual reach coexistence 

We will calculate individual mean coexistence using reach criteria

```{r}

indiv_species_reach_coex <- 
  IndivSpec_reach_richness(df.TS.TE = df_TS_TE_faurby,
                           df.occ = df_occ_faurby2,
                           time.slice = 0.1, 
                           round.digits = 1,
                           species = "species",
                           TS = "TS",
                           TE = "TE",
                           Max.age = "Max.age",
                           Min.age = "Min.age",
                           lat = "lat",
                           lon = "lng")

```

plotting individual species coexistence with reach criteria

```{r}

indiv_species_reach_coex2 <- 
  indiv_species_reach_coex |> 
  mutate(time.slice = as.numeric(time.slice))

indiv_species_reach_coex2 |> 
  filter(species == "Ursus" | species == "Simocyon" | species == "Plionarctos" | species == "Vormela" | species == "Hesperocyon") |> 
  ggplot(aes(x = time.slice, y = n.coexistence, color = species, group = species)) +  # Line thickness for better visibility
  geom_smooth(se = TRUE, method = "loess", size = 1) +
  facet_wrap(~species) +
  labs(title = "",
       x = "Time",
       y = "mean coexistence") +
  theme_minimal()  +
  scale_x_continuous(breaks = seq(max(indiv_species_site_coex2$time.slice), 0, by = -10)) +
  scale_x_reverse() +
  theme(legend.position = "none")
```

